Accelerating Verified-Compiler Development with a Verified Rewriting Engine

### 2022-05-03 papers/source/rewriting/popl2022/rewriting.tex
Add more final versions of 2022 ITP papers

\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
\RequirePackage{luatex85}
\PassOptionsToPackage{dvipsnames}{xcolor}
\RequirePackage{amssymb}
% order matters, see https://tex.stackexchange.com/a/83051/2066
\RequirePackage{varioref}
\PassOptionsToPackage{nameinlink}{cleverref}%[2012/02/15]% v0.18.4;
\documentclass[a4paper,USenglish,cleveref,autoref,thm-restate]{lipics-v2021}
\usepackage{hyperref}
%This is a template for producing LIPIcs articles.
%See lipics-v2021-authors-guidelines.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling autoref support, use "autoref"
%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
%for enabling thm-restate support, use "thm-restate"
%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
%for producing a PDF according the PDF/A standard, add "pdfa"

%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\bibliographystyle{plainurl}% the mandatory bibstyle

%\usepackage{wrapfig}% forbidden in ITP
\setlength{\intextsep}{0pt} % don't add space above nor below
%\usepackage{subfig}
%\newcommand{\subfigureautorefname}{\figureautorefname}

\nolinenumbers% needef for arXiv
\usepackage[frozencache]{minted}
\usemintedstyle{bw}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{comment}
\usepackage[all]{xy}
\usepackage{stmaryrd}
\usepackage[utf8]{inputenc}
\usepackage{textgreek}
\usepackage{newunicodechar}
\usepackage{usebib}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{bcprules}
\usepackage{multicol}
\usepackage{apptools}
\usepackage{placeins}
\usepackage{adjustbox}
\usepackage{xparse}
\usepgfplotslibrary{units}
\usepgfplotslibrary{external}
\tikzexternalize
\newunicodechar{→}{\ensuremath{\to}}
\newunicodechar{λ}{\ensuremath{\lambda}}
\newunicodechar{∀}{\ensuremath{\forall}}

@@

\provideboolean{tobesplit}

\ifthenelse{\boolean{tobesplit}}{%
  \AtBeginDocument{\renewcommand{\appendixautorefname}{Online Appendix}}%
}{}
%https://tex.stackexchange.com/a/149897/2066
\makeatletter
\patchcmd{\hyper@makecurrent}{%
    \ifx\Hy@param\Hy@chapterstring
        \let\Hy@param\Hy@chapapp
    \fi
}{%
    \iftoggle{inappendix}{%true-branch
        % list the names of all sectioning counters here
        \@checkappendixparam{chapter}%
        \@checkappendixparam{section}%
        \@checkappendixparam{subsection}%
        \@checkappendixparam{subsubsection}%
        \@checkappendixparam{paragraph}%
        \@checkappendixparam{subparagraph}%
    }{}%
}{}{\show\hyper@makecurrent\errmessage{failed to patch}}

\newcommand*{\@checkappendixparam}[1]{%
    \def\@checkappendixparamtmp{#1}%
    \ifx\Hy@param\@checkappendixparamtmp
        \let\Hy@param\Hy@appendixstring
    \fi
}
\makeatletter

\newtoggle{inappendix}
\togglefalse{inappendix}

\apptocmd{\appendix}{\toggletrue{inappendix}\renewcommand{\appendixautorefname}{Appendix}}{}{\show\appendix\errmessage{failed to patch}}
%\apptocmd{\subappendices}{\toggletrue{inappendix}}{}{\show\subappendices\errmessage{failed to patch}}
% end appendix autoref patch

@@

\newbibfield{author}
\newbibfield{title}
\newbibfield{citet}
\bibinput{rewriting-lower}
\newcommand{\citetitle}[1]{\emph{\usebibentry{#1}{title}}\nocite{#1}}
\newcommand{\citet}[1]{\usebibentry{#1}{citet}~\cite{#1}}
\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO: #1}}}
\newcommand{\coqbug}[1]{\href{https://github.com/coq/coq/issues/#1}{\##1}}
\newcommand{\coqpr}[1]{\href{https://github.com/coq/coq/pull/#1}{\##1}}
\newcommand{\coqissue}[1]{\href{https://github.com/coq/coq/issues/#1}{\##1}}
\NewDocumentCommand{\githublink}{m m o}{%
  \IfNoValueTF{#3}%
    {\href{https://github.com/#1/#2}{\texttt{#1/#2}}}%
    {\href{https://github.com/#1/#2/tree/#3}{\texttt{#1/#2@#3}}}%
}
\NewDocumentCommand{\githublinkurl}{m m o}{%
  \IfNoValueTF{#3}%
    {\url{https://github.com/#1/#2}}%
    {\url{https://github.com/#1/#2/tree/#3}}%
}
\newcommand{\Rtac}{\ensuremath{\mathcal{R}_\text{\emph{tac}}}}
\newcommand{\Ltac}{\ensuremath{\mathcal{L}_\text{\emph{tac}}}}
\newcommand{\tacvmcompute}{\mintinline{coq}{vm_compute}}
\newcommand{\tacnativecompute}{\mintinline{coq}{native_compute}}
\newcommand{\taccbv}{\mintinline{coq}{cbv}}
\newcommand{\taclazy}{\mintinline{coq}{lazy}}
\newcommand{\tacsimpl}{\mintinline{coq}{simpl}}
\newcommand{\taccbn}{\textbf{\texttt{cbn}}}
\newcommand{\defeq}{=}
\makeatletter
\newcommand{\letin}[1][{\ensuremath{\cdots}}{\ensuremath{\cdots}}]{%
  \texttt{let }\@firstoftwo#1\texttt{ in }\@secondoftwo#1
}
\newcommand{\einput}[1]{\@@input #1 \space}
\newcommand{\beginTikzpictureStamped}[2][]{%
  {%
    \everyeof{\noexpand}% IDK why \noexpand is the magic one, but I got it from http://mirrors.ibiblio.org/CTAN/macros/latex/contrib/oberdiek/catchfile.pdf
    \long\xdef\@tikzstamp{#2}%
  }%
  \def\@dobegintikzpicture{\begin{tikzpicture}[#1]}%
  \expandafter\@dobegintikzpicture\expandafter\def\expandafter\tikzstamp\expandafter{\@tikzstamp}%
}
\newcommand{\TikzpictureStamped}[3][]{\beginTikzpictureStamped[#1]{#2}#3\end{tikzpicture}}%
\makeatother
\newcommand{\asserteq}[2]{\ifthenelse{\equal{#1}{#2}}{}{\GenericError{}{Not equal: \detokenize{#1} != \detokenize{#2}}{}{}}}

\makeatletter
\AtAppendix{%
  \newwrite\@appendixpagefile
  \immediate\openout\@appendixpagefile=\jobname.appendix.startpage\relax
  \immediate\write\@appendixpagefile{\detokenize{\def\appendixstartpage}{\thepage}}%
}
\newcommand{\startsupplement}{%
  \clearpage
  \newwrite\@supplementpagefile
  \immediate\openout\@supplementpagefile=\jobname.supplement.startpage\relax
  \immediate\write\@supplementpagefile{\detokenize{\def\supplementstartpage}{\thepage}}%
  \newwrite\@supplementonlypagefile
  \immediate\openout\@supplementonlypagefile=\jobname.supplement.onlystartpage\relax
  \immediate\write\@supplementonlypagefile{\thepage}%
}%
\AtEndDocument{%
  \newwrite\@endsupplementonlypagefile
  \immediate\openout\@endsupplementonlypagefile=\jobname.supplement.onlyendpage\relax
  \immediate\write\@endsupplementonlypagefile{\thepage}%
}
\makeatother

\title{Accelerating Verified-Compiler Development with a Verified Rewriting Engine}

%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line

\author{Jason Gross}{CSAIL, Massachusetts Institute of Technology, 77 Massachusetts Ave., Cambridge, MA 02139, USA \and MIRI, USA \and \url{https://jasongross.github.io/} }{jgross@mit.edu}{https://orcid.org/0000-0002-9427-4891}{}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty. Please provide at least the name of the affiliation and the country. The full address is optional. Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.

\author{Andres Erbsen}{CSAIL, Massachusetts Institute of Technology, 77 Massachusetts Ave., Cambridge, MA 02139, USA \and \url{https://andres.systems/} }{andreser@mit.edu}{}{}

\author{Jade Philipoom%\footnote{now with Google}
}{CSAIL, Massachusetts Institute of Technology, 77 Massachusetts Ave., Cambridge, MA 02139, USA \and Google}{jade.philipoom@gmail.com}{}{}
% should we \footnote{now with Google}?

\author{Miraya Poddar-Agrawal}{Reed College, 3203 SE Woodstock Blvd, Portland, OR 97202, USA% \and \url{https://github.com/miraya-pa}
}{ragrawal@reed.edu}{https://orcid.org/0000-0001-7617-9180}{}

\author{Adam Chlipala}{CSAIL, Massachusetts Institute of Technology, 77 Massachusetts Ave., Cambridge, MA 02139, USA \and \url{http://adam.chlipala.net/} }{adamc@csail.mit.edu}{https://orcid.org/0000-0001-7085-9417}{%[funding]
}

\authorrunning{J. Gross and A. Erbsen and J. Philipoom and M. Poddar-Agrawal and A. Chlipala} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

\Copyright{Jason Gross \& Andres Erbsen \& Jade Philipoom \& Miraya Poddar-Agrawal \& Adam Chlipala} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10003752.10003790.10002990</concept_id>
       <concept_desc>Theory of computation~Logic and verification</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10003752.10003790.10003798</concept_id>
       <concept_desc>Theory of computation~Equational logic and rewriting</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10011006.10011041</concept_id>
       <concept_desc>Software and its engineering~Compilers</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10011006.10011041.10011046</concept_id>
       <concept_desc>Software and its engineering~Translator writing systems and compiler generators</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[300]{Theory of computation~Logic and verification}
\ccsdesc[500]{Theory of computation~Equational logic and rewriting}
\ccsdesc[300]{Software and its engineering~Compilers}
\ccsdesc[300]{Software and its engineering~Translator writing systems and compiler generators}

\keywords{compiler verification, rewriting engines, cryptography} %TODO mandatory; please add comma-separated list of keywords

%\category{} %optional, e.g. invited paper

\ifthenelse{\boolean{tobesplit}}{%
%  \relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%  \relatedversiondetails%[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]
%      {Full Version}{https://arxiv.org/abs/2205.00862} %linktext and cite are optional
}{}

\supplement{%
%  \url{https://jasongross.github.io/papers/2022-rewriting-itp-supplement.pdf},
  \githublinkurl{mit-plv}{rewriter}[ITP-2022-perf-data],
  \githublinkurl{mit-plv}{fiat-crypto}[perf-testing-data-ITP-2022-rewriting]%
}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
%\supplementdetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93, subcategory={Description, Subcategory}, swhid={Software Heritage Identifier}]{General Classification (e.g. Software, Dataset, Model, ...)}{URL to related version} %linktext, cite, and subcategory are optional

\funding{%
  This work was supported in part by a Google Research Award, National Science Foundation grants CCF-1253229, CCF-1512611, and CCF-1521584, and the National Science Foundation Graduate Research Fellowship under Grant Nos.\ 1122374 and 1745302.
  Any opinion, findings, and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the National Science Foundation.%
}
%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.

%\acknowledgements{I want to thank \dots}%optional

%\nolinenumbers %uncomment to disable line numbering

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{June Andronick and Leonardo de Moura}
\EventNoEds{2}
\EventLongTitle{13th International Conference on Interactive Theorem Proving (ITP 2022)}
\EventShortTitle{ITP 2022}
\EventAcronym{ITP}
\EventYear{2022}
\EventDate{August 7--10, 2022}
\EventLocation{Haifa, Israel}
\EventLogo{}
\SeriesVolume{237}
\ArticleNo{5}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

%TODO mandatory: add short abstract of the document
\begin{abstract}
  Compilers are a prime target for formal verification, since compiler bugs invalidate higher-level correctness guarantees, but compiler changes may become more labor-intensive to implement, if they must come with proof patches.
  One appealing approach is to present compilers as sets of algebraic rewrite rules, which a generic engine can apply efficiently.
  Now each rewrite rule can be proved separately, with no need to revisit past proofs for other parts of the compiler.
  We present the first realization of this idea, in the form of a framework for the Coq proof assistant.
  Our new Coq command takes normal proved theorems and combines them automatically into fast compilers with proofs.
  We applied our framework to improve the Fiat Cryptography toolchain for generating cryptographic arithmetic, producing an extracted command-line compiler that is about 1000$\times$ faster while actually featuring simpler compiler-specific proofs.
\end{abstract}

@@

Formally verified compilers like CompCert~\cite{Compcert} and CakeML~\cite{CakeML} are success stories for proof assistants, helping close a trust gap for one of the most important categories of software infrastructure.
A popular compiler cannot afford to stay still; developers will add new backends, new language features, and better optimizations.
Proofs must be adjusted as these improvements arrive.
It makes sense that the author of a new piece of compiler code must prove its correctness, but ideally there would be no need to revisit old proofs.
There has been limited work, though, on avoiding that kind of coupling.
\citet{xcert} demonstrated a streamlined way to extend CompCert with new verified optimizations driven by dataflow analysis, but we are not aware of past work that supports easy extension for compilers from functional languages to C code.
We present our work targeting that style.

@@

In this paper, we present \textbf{the first proven-correct compiler-builder toolkit parameterized on rewrite rules}.
Arbitrary sets of Coq theorems (quantified equalities) can be assembled by a single new Coq command into an extraction-ready verified compiler.
We did not need to extend the trusted code base, so our compiler compiler need not be trusted.
We achieve both good performance of compiler runs and good performance of generated code, via addressing a number of scale-up challenges vs.\ past work.
%including efficient handling of intermediate code with thousands of nested binders and critical subterm sharing.

We evaluate our toolkit by replacing a key component of Fiat Cryptography~\cite{FiatCryptoSP19}, a Coq library that generates code for big-integer modular arithmetic at the heart of elliptic-curve-cryptography algorithms.
Routines generated (with proof) with Fiat Cryptography now ship with all major Web browsers and all major mobile operating systems.
With our improved compiler architecture, it became easy to add two new backends and a variety of new supported source-code features, and we were easily able to try out new optimizations.
%and later replace unsound optimizations with equivalent sound optimizations.

Replacing Fiat Cryptography's original compiler with the compiler generated by our toolkit has two additional benefits.
Fiat Cryptography was previously only used successfully to build C code for two of the three most widely used curves (P-256 and Curve25519).
Our prior version's execution timed out trying to compile code for the third most widely used curve (P-384).
Using our new toolkit has made it possible to generate compiler-synthesized code for P-384 while generating completely identical code for the primes handled by the previous version, about 1000$\times$ more quickly.
Additionally, Fiat Cryptography previously required source code to be written in continuation-passing style, and our compiler has enabled a direct-style approach, which pays off in simplifying theorem statements and proofs.

@@

\subsection{Related Work}\label{sec:related-one}
% POPL 2022 review TODO:
% Section 1.1, it may be a little early at this point to discuss the related work, as I don't think it is quite clear yet exactly what problem this paper solves.
% From the first 1.5 pages, I understand that the Fiat Cryptography compiler has been improved, but it is not clear yet how this is achieved and what are the merits or demerits of the approach presented in this paper.

Assume our mission is to take libraries of purely functional combinators, apply them to compile-time parameters, and compile the results down to lean C code.
Furthermore, we ask for machine-checked proofs that the C programs preserve the behavior of the higher-order functional programs we started with.
What good ideas from the literature can we build on?

>>>>

\RequirePackage{luatex85}
\PassOptionsToPackage{dvipsnames}{xcolor}
%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\RequirePackage{amssymb}
% order matters, see https://tex.stackexchange.com/a/83051/2066
\RequirePackage{varioref}
\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
\usepackage{hyperref}
\usepackage[nameinlink]{cleveref}[2012/02/15]% v0.18.4;
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2018}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
%\usepackage{booktabs}   %% For formal tables:
%                        %% http://ctan.org/pkg/booktabs
%\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
%                        %% http://ctan.org/pkg/subcaption

\usepackage{wrapfig}
\setlength{\intextsep}{0pt} % don't add space above nor below
\usepackage{subfig}
\newcommand{\subfigureautorefname}{\figureautorefname}
%% http://ctan.org/pkg/subcaption

\usepackage{minted}
\usemintedstyle{bw}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{comment}
\usepackage[all]{xy}
\usepackage{stmaryrd}
\usepackage[utf8]{inputenc}
\usepackage{textgreek}
\usepackage{newunicodechar}
\usepackage{usebib}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{bcprules}
\usepackage{multicol}
\usepackage{apptools}
\usepackage{adjustbox}
\usepgfplotslibrary{units}
\usepgfplotslibrary{external}
\tikzexternalize
\newunicodechar{→}{\ensuremath{\to}}
\newunicodechar{λ}{\ensuremath{\lambda}}
\newunicodechar{∀}{\ensuremath{\forall}}

@@

@@

% https://tex.stackexchange.com/a/183682/2066
\makeatletter

% define a macro \Autoref to allow multiple references to be passed to \autoref
\newcommand\Autoref[1]{\@first@ref#1,@}
\def\@throw@dot#1.#2@{#1}% discard everything after the dot
\def\@set@refname#1{%    % set \@refname to autoefname+s using \getrefbykeydefault
    \edef\@tmp{\getrefbykeydefault{#1}{anchor}{}}%
    \xdef\@tmp{\expandafter\@throw@dot\@tmp.@}%
    \ltx@IfUndefined{\@tmp autorefnameplural}%
         {\def\@refname{\@nameuse{\@tmp autorefname}s}}%
         {\def\@refname{\@nameuse{\@tmp autorefnameplural}}}%
}
\def\@first@ref#1,#2{%
  \ifx#2@\autoref{#1}\let\@nextref\@gobble% only one ref, revert to normal \autoref
  \else%
    \@set@refname{#1}%  set \@refname to autoref name
    \@refname~\ref{#1}% add autoefname and first reference
    \let\@nextref\@next@ref% push processing to \@next@ref
  \fi%
  \@nextref#2%
}
\def\@next@ref#1,#2{%
   \ifx#2@ and~\ref{#1}\let\@nextref\@gobble% at end: print and+\ref and stop
   \else, \ref{#1}% print  ,+\ref and continue
   \fi%
   \@nextref#2%
}

\makeatother

\newbibfield{author}
\newbibfield{title}
\newbibfield{citet}
\bibinput{rewriting-lower}
\newcommand{\citetitle}[1]{\emph{\usebibentry{#1}{title}}\nocite{#1}}
\newcommand{\customcitet}[1]{\usebibentry{#1}{citet}~\cite{#1}}
\newcommand{\todo}[1]{\textbf{TODO: #1}}
\newcommand{\coqbug}[1]{\href{https://github.com/coq/coq/issues/#1}{\##1}}
\newcommand{\coqpr}[1]{\href{https://github.com/coq/coq/pull/#1}{\##1}}
\newcommand{\coqissue}[1]{\href{https://github.com/coq/coq/issues/#1}{\##1}}
\newcommand{\Rtac}{\ensuremath{\mathcal{R}_\text{\emph{tac}}}}
\newcommand{\Ltac}{\ensuremath{\mathcal{L}_\text{\emph{tac}}}}
\newcommand{\tacvmcompute}{\mintinline{coq}{vm_compute}}
\newcommand{\tacnativecompute}{\mintinline{coq}{native_compute}}
\newcommand{\taccbv}{\mintinline{coq}{cbv}}
\newcommand{\taclazy}{\mintinline{coq}{lazy}}
\newcommand{\tacsimpl}{\mintinline{coq}{simpl}}
\newcommand{\taccbn}{\textbf{\texttt{cbn}}}
\newcommand{\defeq}{=}
\makeatletter
\newcommand{\letin}[1][{\ensuremath{\cdots}}{\ensuremath{\cdots}}]{%
  \texttt{let }\@firstoftwo#1\texttt{ in }\@secondoftwo#1
}
\newcommand{\einput}[1]{\@@input #1 \space}
\newcommand{\beginTikzpictureStamped}[2][]{%
  {%
    \everyeof{\noexpand}% IDK why \noexpand is the magic one, but I got it from http://mirrors.ibiblio.org/CTAN/macros/latex/contrib/oberdiek/catchfile.pdf
    \long\xdef\@tikzstamp{#2}%
  }%
  \def\@dobegintikzpicture{\begin{tikzpicture}[#1]}%
  \expandafter\@dobegintikzpicture\expandafter\def\expandafter\tikzstamp\expandafter{\@tikzstamp}%
}
\newcommand{\TikzpictureStamped}[3][]{\beginTikzpictureStamped[#1]{#2}#3\end{tikzpicture}}%
\makeatother
\newcommand{\asserteq}[2]{\ifthenelse{\equal{#1}{#2}}{}{\GenericError{}{Not equal: \detokenize{#1} != \detokenize{#2}}{}{}}}

\makeatletter
\AtAppendix{%
  \newwrite\@appendixpagefile
  \immediate\openout\@appendixpagefile=\jobname.appendix.startpage\relax
  \immediate\write\@appendixpagefile{\detokenize{\def\appendixstartpage}{\thepage}}%
}
\newcommand{\supplement}{%
  \clearpage
  \newwrite\@supplementpagefile
  \immediate\openout\@supplementpagefile=\jobname.supplement.startpage\relax
  \immediate\write\@supplementpagefile{\detokenize{\def\supplementstartpage}{\thepage}}%
  \newwrite\@supplementonlypagefile
  \immediate\openout\@supplementonlypagefile=\jobname.supplement.onlystartpage\relax
  \immediate\write\@supplementonlypagefile{\thepage}%
}%
\AtEndDocument{%
  \newwrite\@endsupplementonlypagefile
  \immediate\openout\@endsupplementonlypagefile=\jobname.supplement.onlyendpage\relax
  \immediate\write\@endsupplementonlypagefile{\thepage}%
}
\makeatother

\begin{document}
%% Title information
%\title[Short Title]{Full Title}         %% [Short Title] is optional;
%                                        %% when present, will be used in
%                                        %% header instead of Full Title.
%\titlenote{with title note}             %% \titlenote is optional;
%                                        %% can be repeated if necessary;
%                                        %% contents suppressed with 'anonymous'
%\subtitle{Subtitle}                     %% \subtitle is optional
%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
%                                        %% can be repeated if necessary;
%                                        %% contents suppressed with 'anonymous'
\title{Accelerating Verified-Compiler Development with a Verified Rewriting Engine}

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{Jason Gross}
%\authornote{with author1 note}          %% \authornote is optional;
%                                        %% can be repeated if necessary
\orcid{0000-0002-9427-4891}
\affiliation{
%  \position{Position1}
  \department{CSAIL}              %% \department is recommended
  \institution{Massachusetts Institute of Technology}            %% \institution is required
  \streetaddress{77 Massachusetts Avenue}
  \city{Cambridge}
  \state{MA}
  \postcode{02139}
  \country{USA}                    %% \country is recommended
}
\email{jgross@mit.edu}          %% \email is recommended

\author{Andres Erbsen}
%\authornote{with author1 note}          %% \authornote is optional;
%                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
%  \position{Position1}
  \department{CSAIL}              %% \department is recommended
  \institution{Massachusetts Institute of Technology}            %% \institution is required
  \streetaddress{77 Massachusetts Avenue}
  \city{Cambridge}
  \state{MA}
  \postcode{02139}
  \country{USA}                    %% \country is recommended
}
\email{andreser@mit.edu}          %% \email is recommended

\author{Rajashree Agrawal}
%\authornote{with author1 note}          %% \authornote is optional;
%                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
%  \position{Position1}
%  \department{}              %% \department is recommended
  \institution{Reed College}            %% \institution is required
  \streetaddress{3203 SE Woodstock Blvd}
  \city{Portland}
  \state{OR}
  \postcode{97202}
  \country{USA}                    %% \country is recommended
}
\email{ragrawal@reed.edu}          %% \email is recommended

\author{Jade Philipoom}
\authornote{now with Google}          %% \authornote is optional;
                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
%  \position{Position1}
  \department{CSAIL}              %% \department is recommended
  \institution{Massachusetts Institute of Technology}            %% \institution is required
  \streetaddress{77 Massachusetts Avenue}
  \city{Cambridge}
  \state{MA}
  \postcode{02139}
  \country{USA}                    %% \country is recommended
}
\email{jadep@mit.edu}          %% \email is recommended

\author{Adam Chlipala}
%\authornote{with author1 note}          %% \authornote is optional;
%                                        %% can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
%  \position{Position1}
  \department{CSAIL}              %% \department is recommended
  \institution{Massachusetts Institute of Technology}            %% \institution is required
  \streetaddress{77 Massachusetts Avenue}
  \city{Cambridge}
  \state{MA}
  \postcode{02139}
  \country{USA}                    %% \country is recommended
}
\email{adamc@csail.mit.edu}          %% \email is recommended

%
%\authorrunning{J.\ Gross et al.\@}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  Compilers are a prime target for formal verification, since compiler bugs invalidate higher-level correctness guarantees, but compiler changes may become more labor-intensive to implement, if they must come with proof patches.
  One appealing approach is to present compilers as sets of algebraic rewrite rules, which a generic engine can apply efficiently.
  Now each rewrite rule can be proved separately, with no need to revisit past proofs for other parts of the compiler.
  We present the first realization of this idea, in the form of a framework for the Coq proof assistant.
  Our new Coq command takes normal proved theorems and combines them automatically into fast compilers with proofs.
  We applied our framework to improve the Fiat Cryptography toolchain for generating cryptographic arithmetic, producing an extracted command-line compiler that is about 1000$\times$ faster while actually featuring simpler compiler-specific proofs.
\end{abstract}

%\keywords{First keyword  \and Second keyword \and Another keyword.}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
%\keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

@@

Formally verified compilers like CompCert~\cite{Compcert} and CakeML~\cite{CakeML} are success stories for proof assistants, helping close a trust gap for one of the most important categories of software infrastructure.
A popular compiler cannot afford to stay still; developers will add new backends, new language features, and better optimizations.
Proofs must be adjusted as these improvements arrive.
It makes sense that the author of a new piece of compiler code must prove its correctness, but ideally there would be no need to revisit old proofs.
There has been limited work, though, on avoiding that kind of coupling.
\citet{xcert} demonstrated a streamlined way to extend CompCert with new verified optimizations driven by dataflow analysis, but we are not aware of past work that supports easy extension for compilers from functional languages to C code.
We present our new work targeting that sort of compilation.

@@

In this paper, we present \textbf{the first proven-correct compiler-builder toolkit parameterized on rewrite rules}.
Arbitrary sets of Coq theorems, proving quantified equalities, can be assembled by a single new Coq command into an extraction-ready verified compiler.
We did not need to extend the trusted code base, so our compiler compiler need not be trusted in any way.
We achieve both good performance of compiler runs and good performance of generated code, via addressing a number of scale-up challenges vs.\ past work, including efficient handling of intermediate code with thousands of nested variable binders and critical subterm sharing.

We evaluate our builder toolkit by replacing a key component of Fiat Cryptography, a Coq library that generates code for big-integer modular arithmetic at the heart of elliptic-curve-cryptography algorithms.
This domain-specific compiler has been adopted, for instance, in the Chrome Web browser, such that the majority of all HTTPS connections from browsers are now initiated using code generated (with proof) by Fiat Cryptography.
With our improved compiler architecture, it became easy to add two new backends and a variety of new supported source-code features, and we were easily able to try out new optimizations and later replace unsound optimizations with equivalent sound optimizations.

Replacing Fiat Cryptography's original compiler with the compiler generated by our toolkit has two additional benefits.
Fiat Cryptography was previously only used successfully to build C code for the two most widely used curves (P-256 and Curve25519).
Their execution timed out trying to compile code for the third most widely used curve (P-384).
Using our toolkit has made it possible to generate compiler-synthesized code for P-384 while generating completely identical code for the primes handled by the previous version, about 1000$\times$ more quickly.
Additionally, Fiat Cryptography previously required source code to be written in continuation-passing style, and our compiler has enabled a direct-style approach, which pays off in simplifying the theorem statement and proof for every arithmetic combinator.

@@

\subsection{Related Work, Take One}\label{sec:related-one}
% POPL 2022 review TODO:
% Section 1.1, it may be a little early at this point to discuss the related work, as I don't think it is quite clear yet exactly what problem this paper solves.
% From the first 1.5 pages, I understand that the Fiat Cryptography compiler has been improved, but it is not clear yet how this is achieved and what are the merits or demerits of the approach presented in this paper.

Assume our mission is to take libraries of purely functional combinators, apply them to compile-time parameters, and compile the results down to lean C code.
Furthermore, we ask for machine-checked proofs that the C programs preserve the input-output behavior of the higher-order functional programs we started with.
What good ideas from the literature can we build on?
